---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# RCTRecruit

<!-- badges: start -->
[![R-CMD-check](https://github.com/imalagaris/RCTRecruit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/imalagaris/RCTRecruit/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/imalagaris/RCTRecruit/graph/badge.svg)](https://app.codecov.io/gh/imalagaris/RCTRecruit)
<!-- badges: end -->

## Installation
Accurate prediction of subject recruitment for randomized clinical
    trials (RCT) remains an ongoing challenge. Many previous prediction models rely
    on parametric assumptions. We present functions for non-parametric RCT
    recruitment prediction under several scenarios.  


You can install the development version of RCTRecruit from 
[GitHub](https://github.com/imalagaris/RCTRecruit) with:

``` r
devtools::install_github("imalagaris/RCTRecruit")
```

## Example

```{r example}
library(RCTRecruit)
LoadData(gripsYR1, ScreenDt, Enrolled)
Time2Nsubjects()
GetDistance(gripsYR2Weekly$enrolled)
res <- GetWeekPredCI()
plot(res)
```


